lifespan_eda

### 2023-12-03 lifespan_eda.Rmd
clean up nas

#drop repeat scans
cn.df.unique <- cn.df %>%
  drop_na(c(age_days, participant, study)) %>%
  arrange(age_days) %>%
  group_by(participant) %>%
  slice(1) %>%
  ungroup()

@@

@@

```{r}
cn.df.site <- cn.df.unique %>% 
  dplyr::filter(site != "NA") %>%
  mutate(study = droplevels(study),
         site = gsub(" ", "_", trimws(site)))
n_unique(cn.df.site$participant)
n_unique(cn.df.site$participant) == nrow(cn.df.site)

@@

```{r}
#clean up site NAs
cn.df.imp <- cn.df.unique %>%
  mutate(site = trimws(as.character(site)),
         study = trimws(study)) %>%
  mutate(site = gsub(" ", "_", ifelse(is.na(site), study, site)))

>>>>

#drop repeat scans
cn.df.unique <- cn.df %>%
  drop_na(c(age_days, participant, study)) %>%
  arrange(age_days) %>%
  group_by(participant) %>%
  slice(1) %>%
  ungroup() %>%
  mutate(sexMale = as.numeric(case_when(sex == "Male" ~ 1,
                                        sex == "Female" ~ 0,
                                        TRUE ~ NA)),
         sex.age = sexMale*age_days,
         log_age = log(age_days, base=10))

@@

```{r}
skim(cn.df.unique)
levels(cn.df.unique$study)
```

## Visualize Features
plot global brain features for basic qc - check again post-combat

```{r}
ggplot(cn.df.unique) +
  geom_point(aes(x=age_days, y=GMV, color=sex))

ggplot(cn.df.unique) +
  geom_point(aes(x=age_days, y=sGMV, color=sex))

ggplot(cn.df.unique) +
  geom_point(aes(x=age_days, y=WMV, color=sex))

ggplot(cn.df.unique) +
  geom_point(aes(x=age_days, y=Ventricles, color=sex))
```

log-scaled
```{r}
ggplot(cn.df.unique) +
  geom_point(aes(x=log_age, y=GMV, color=sex))

ggplot(cn.df.unique) +
  geom_point(aes(x=log_age, y=sGMV, color=sex))

ggplot(cn.df.unique) +
  geom_point(aes(x=log_age, y=WMV, color=sex))

ggplot(cn.df.unique) +
  geom_point(aes(x=log_age, y=Ventricles, color=sex))
```

@@

```{r}
cn.df.site <- cn.df.unique %>% 
  dplyr::filter(site != "NA") %>%
  mutate(study = droplevels(study),
         site = gsub(" ", "_", trimws(site))) %>%
  na.omit()
n_unique(cn.df.site$participant)
n_unique(cn.df.site$participant) == nrow(cn.df.site)

@@

```{r}
#clean up site NAs
cn.df.imp <- cn.df.unique %>%
  mutate(site = trimws(as.character(site)),
         study = trimws(study)) %>%
  mutate(site = gsub(" ", "_", ifelse(is.na(site), study, site))) %>%
  na.omit()
